%
% r-3
%
\newpage

\begin{fullwidth}

	\thispagestyle{empty}
	\setlength{\parindent}{0pt}
	\setlength{\parskip}{\baselineskip}

  % \vskip 8em

	{
		\LARGE%
		\par%
			\smallcaps{François Briatte}%
			%
	}\\[0.167\textheight]
		
	\vskip 12em
		
	{\normalsize\smallcaps{this is stata, or}}\\[.5\baselineskip]
	{\huge\allcaps{Stata for social scientists}}\\[\baselineskip]

	\vfill

	\index{Acknowledgements}%
	%
  \smallcaps{Acknowledgements} go first and foremost to Ivaylo Petev, with whom I co-taught many of the classes for which this guide was written, and to the students who offered their invaluable feedback on earlier draft versions, with a special mention to Laura Führer. I have also benefitted from the friendly advice of Joël Gombin, Emiliano Grossman, Filip Kostelka, Sarah McLaughlin, Dawn Teele, Vincent Tiberj and Hyungsoo Woo. Final thanks to the \textsc{psia} and \textsc{glm} staff at Sciences~Po for their excellent support. All mistakes are mine alone.%

  \smallcaps{The Stata software} and its logo are registered trademarks of StataCorp~LP. The cover image is derived from a screenshot of Zack Snyder's \emph{300} (Warner Brothers, 2007). Please refer to the code and text for other attributions.%

  \smallcaps{Please do not cite this draft} and send all feedback to \href{mailto:f.briatte@ed.ac.uk}{f.briatte@ed.ac.uk}, or use the development repository to report an issue: \url{https://github.com/briatte/srqm} -- thanks in advance for your help!%

\end{fullwidth}

\cleardoublepage
