%
% 1
%
\chapter{Introduction}%
	%
	%
	\label{ch:introduction}%
	\index{Course!Introduction}%
	%
	%

	\newthought{Dear reader,}\\[1em]%
	%
	\newthought{You probably find yourself reading this} because you enrolled in Statistical Reasoning and Quantitative Methods (\SRQM), a postgraduate introduction to statistics that Ivaylo Petev and myself have been teaching at Sciences~Po in Paris since 2010. Welcome!%

	This guide is a practical handbook to read while taking the course. The rest of the teaching material, which includes the syllabus and essential files to be used in and outside class, can be downloaded from the course webpage at the following address:\\[1em]%

		\url{http://f.briatte.org/teaching/quanti/}\\[1em]%
		
	\newthought{Start reading below} to learn about the course basics. You will be expected to have read this section in your first week of class. In particular, you should make sure that you have \textbf{run the course setup} on your computer, and that you know how to run it again if necessary.%

	%
	%
	\minitoc
	%
	\newpage
	%
	%

	%
	% 1.1
	%
	\include{1A_topic}%
	%
	% 1.2
	%
	\include{1B_stata}%
	%
	% 1.3
	%
	\include{1C_setup}%
	%
	%
	
	\subsection{Last introductory words}

	\newthought{When you are done with your work}, just quit Stata like you would quit any other program.%
		%
		%
	
	At that stage, Stata will close any opened logs and all unsaved results will be lost—which is fine, because you should \hlred{\textbf{never save any changes to the original dataset}}: keep the teaching material intact at all times by systematically discarding all changes. Just make sure to save your do-files if you have not yet done so.%
		%
		%

	The nerdy way to quit Stata is to quit from the command line. The \cmd{exit} command with the \opt{clear}{exit} will erase any data in memory and quit:%
		%
		%

		\begin{docspec}
			* Close all open logs.\\%
			log close \_all\\[1em]%
			%
			* Enjoy your day.\\%
			exit, clear
		\end{docspec}
		%
		%
